### 2021-03-01 writeups/FuelsWeatherFirebehaviorFigures.Rmd
begin final draft of figures

---
title: Fuels, fire weather, and Rx fire behavior in North Dakota
subtitle: Finalizing figures and results
date: '`r Sys.Date()`'
author: DAM
output:
  html_document:
    keep_md: false
    toc: true
    toc_float: true
    theme: united
    highlight: tango
---

```{r setup, echo=FALSE, warning=FALSE, message = FALSE, results='hide'}
knitr::opts_chunk$set(message = FALSE, warning=FALSE, 
               echo=FALSE, eval=TRUE, fig.path='figures/', dev=c('png', 'pdf')) 
```

```{r data_loading}
# Packages 
  pacman::p_load(tidyverse, readr, mice, broom.mixed, vegan,
                 lubridate, gridExtra, wesanderson)
set.seed(42)
# Data
  fp = url('https://raw.githubusercontent.com/devanmcg/SpatialFireBehavior/main/data/AnalysisDataKY.csv') 
  AnalysisData <- read_csv(fp) 
# Extra scripts
  source("https://raw.githubusercontent.com/devanmcg/IntroRangeR/master/11_IntroMultivariate/ordinationsGGplot.R")
  source('https://raw.githubusercontent.com/cran/mice/master/R/mipo.R')
```

# Methods

## Regression analysis

Figures generated by the script below and included in this document are [available on GitHub](https://github.com/devanmcg/SpatialFireBehavior/tree/main/writeups/figures), as well.

```{r data_prep, eval=FALSE}
###
### N O T   R U N
###
 AllData <-  
    read_csv(paste0(FilePath, "/data/fromMZ/CompiledData2.csv")) %>%
    filter(location != "OAK") %>%
      mutate(date = as.Date(date, format = "%m/%d/%Y"),
               L = str_remove(location, "REC"), 
               B = str_sub(block, 1,3), 
               Ps = str_replace(pasture, "[.]", ""), 
               Ps = str_sub(Ps, 1,2), 
               patch = str_replace(patch, "[.]", ""),
               y = format(date, "%y")) %>%
        unite("FireCode", c(L,B,Ps,patch,y), sep=".") %>%
    mutate(time = str_remove(MaxTempTime, "[.]+[0-9]"))%>%
    unite(timestamp, c(date, time), sep = " ") %>%
    mutate(timestamp = as.POSIXct(timestamp, format = "%Y-%m-%d %H:%M:%S")) %>%
        select(FireCode, timestamp, plot, array, TC, MaxC, 
               AirTemp, RH, dpC, WindSpeed, 
               LAI, FMC, KgHa) 
# Isolate soil surface temperature (TC 4)
  SoilTemp <-
    filter(AllData, TC == 4) %>%
      select(FireCode, plot, array, MaxC) %>%
        rename(SoilC = MaxC)
# Summarize array-level data 
  DataMeans <- 
    AllData %>%
      filter(TC %in% c('1', '2', '3')) %>% 
      select(-timestamp) %>%
      pivot_longer(cols = c(MaxC:KgHa), 
                   names_to = "var",
                   values_to = "value") %>%
      group_by(FireCode, plot, array, var) %>%
        summarize(Mean = mean(value) ) %>%
      ungroup() %>%
      pivot_wider(names_from = var, 
                  values_from = Mean) 
# Calculate Vapor Pressure Deficit
  DataMeans <- 
    DataMeans %>%
      mutate(e  = 6.11*(10^((7.5*dpC)/(237.3+dpC))), 
             es = 6.11*(10^((7.5*AirTemp)/(237.3+AirTemp))), 
             VPD = es - e) %>%
      select(-e, -es)
# Calculate ROS 
  D = 1   # Distance between thermocouples (m)
  ROS <- 
    AllData %>%
      filter(TC %in% c('1', '2', '3')) %>%
      mutate(timestamp = format(timestamp, "%H:%M:%S"), 
             ArrivalTime = seconds(hms(timestamp)) ) %>%
    select(FireCode, plot, array, ArrivalTime) %>%
    group_by(FireCode, plot, array) %>%
    arrange(ArrivalTime, .by_group = TRUE) %>% 
    mutate(position = order(order(ArrivalTime, decreasing=FALSE)), 
           position = recode(position, "1"="a", "2"="b", "3"="c"), 
           ArrivalTime = as.numeric(ArrivalTime) /60 ) %>%
    spread(position, ArrivalTime)  %>%
    ungroup %>% 
    mutate( theta_rad = atan((2*c - b - a) / (sqrt(3)*(b - a))), 
            ros = case_when(
              a == b ~ (sqrt(3) / 2) / (c - a) , 
              a != b ~  (D*cos(theta_rad) / (b - a) ) 
            )) %>%
    select(-a, -b, -c, -theta_rad)
# Combine, filter into final tibble for analysis 
  AnalysisData  <- 
    full_join(DataMeans, ROS) %>%
              left_join(SoilTemp) %>%
      filter( ros <= 40, 
              MaxC >= 40) %>%
      rename(FuelMoisture = FMC, 
             SoilMaxC = SoilC) %>%
        mutate(FuelMoisture = ifelse(FuelMoisture >= 0, 
                                      FuelMoisture, NA), 
               FuelMoisture = FuelMoisture * 100) %>%
      separate(FireCode, into = c("location", "block", "pasture", 
                           "patch", "year"), 
               remove = F)
###
###
###
```

For analysis, data are filtered to $\textsf{Rate of spread < 40 m/s}$ and $\textsf{Maximum temperature > 40}^\circ \textsf{C}$.

### Imputation of missing values

We used the multiple imputation method in the **mice** package to fill in missing values. 
The method creates 5 datasets, each with different but reasonable values for the missing data based on patterns in existing data, and performs the regression analysis on each. 
These results are pooled into a single set of results.

```{r imp, eval=TRUE}
# Calculate imputed datasets on scaled data
  imp_sc <- AnalysisData %>% 
          select(-LAI, -JD) %>%
          mutate_at(vars(AirTemp:tHa), ~as.numeric(scale(., center=F))) %>%
                    mutate(across(location:array, as.factor)) %>%
                     mice(m=5, seed = 23109, print=F)
# Reduce mids object to tibble for multivariate analysis    
    imp_raw <- 
      complete(imp_sc, 'long') %>%
      as_tibble() %>%
      unite("TreeID", c(location, block, pasture,
                        year, plot, array), sep = ".") %>%
      select(-patch, -.id, -.imp, -FireCode) %>%
      pivot_longer(names_to = "response", 
                   values_to = "values", 
                   -TreeID) %>%
      group_by(TreeID, response) %>%
      summarize(value = median(values)) %>%
      ungroup() %>%
      pivot_wider(names_from = response, 
                  values_from = value) %>%
      separate(TreeID, c("location", "block", "pasture",
                         "year", "plot", "array")) %>% 
      mutate(across(location:array, as.factor))
```

### Model specification

Mixed-effect regression models fit with `lme4::glmer` and constructed as follows:

$$y \approx RH + WS + FM + FL + (1|location/year/plot),~ 
\text{family = gamma, link = log}$$

where

$RH$ = Relative humidity,

$WS$ = Average wind speed,

$FM$ = Fuel moisture (from clipped samples),

$FL$ = Fuel load (t/ha from clipped samples)

Weather variables $RH$ and $WS$ were taken from NDAWN stations for the hour in which the fire behavior observation occurred.

All variables were scaled to a common range prior to analysis, and units for each predictor variable are not reported.

Response variables included

* Maximum temperature (C) 15 cm above soil surface (mean of three thermocouples)
* Maximum temperature (C) at soil surface (single thermocouple)
* Rate of spread (m/s) through three points of a 1 m equilateral triangle

Thermocouple data recorded with the open-source [FeatherFlame](https://diyfirescience.info/portfolio/work2/) system.

### Multivariate analysis

Principal Components Analysis (PCA) on fire behavior responses fit with `vegan::rda`. 
Post-hoc group (location) and gradient (fire weather) analysis with `vegan::envfit`.

An additional composite variable, vapor pressure deficit (VPD), was calculated as

$\text{VPD} = e - e_s$

where

$e = 6.11 \cdot 10^{\frac{7.5 \cdot \text{Dewpoint}}{237.3 + \text{Dewpoint}}}$

and

$e_s = 6.11 \cdot 10^{\frac{7.5 \cdot \text{Air temperature}}{237.3 + \text{Air temperature}}}$

# Results

## Fuel, weather, and fire behavior across locations

```{r data_summary_gg, fig.height=10, fig.width=6}
LongData <- 
  AnalysisData %>%
    select(-LAI, -JD, -MaxWindSpeed) %>%
    pivot_longer(cols = AirTemp:tHa, 
                 names_to = 'var', 
                 values_to = 'value') %>%
    mutate(type = case_when(
                    var %in% c('MaxC', 
                               'SoilMaxC', 
                               'ros') ~ 'Fire behavior', 
                    var %in% c('tHa',
                               'FuelMoisture') ~ 'Fuel', 
                    TRUE ~ 'Weather' ), 
           var = recode(var, 
                        AirTemp = 'Air temperature (°C)', 
                        RH = 'Relative humidity (%)', 
                        dpC = 'Dew point (°C)', 
                        FuelMoisture = 'Fuel moisture (%)', 
                        tHa = 'Fuel load (t/ha)',
                        MaxC = 'Flame temp (°C)', 
                        WindSpeed = 'Wind speed (m/s)', 
                        ros = 'Rate of spread (m/s)', 
                        SoilMaxC = 'Soil surface temp (°C)'), 
           location = recode(location, 
                             H = "Hettinger", 
                             CG = "Central Grasslands"))

wx_d <-
LongData %>%
  filter(type == "Weather") %>%
  ggplot(aes(x = location, y= value,
             fill = location)) + theme_bw(14) +
    geom_violin() +
  labs(y = NULL, x = NULL, 
       title = "Weather") + 
  facet_wrap( ~ var, scales = 'free') +
  scale_fill_manual(name = "Location", 
                    values = wes_palette("Cavalcanti1")[4:5]) + 
  theme(axis.text.x = element_blank(), 
        panel.grid.major.x = element_blank(), 
        legend.position = "none")

fuel_d <-
LongData %>%
  filter(type == "Fuel") %>%
  ggplot(aes(x = location, y= value,
             fill = location)) + theme_bw(14) +
    geom_violin() +
  labs(y = NULL, x = NULL, 
       title = "Fuel") + 
  facet_wrap( ~ var, scales = 'free') +
  scale_fill_manual(name = "Location", 
                    values = wes_palette("Cavalcanti1")[4:5]) + 
  theme(axis.text.x = element_blank(), 
        panel.grid.major.x = element_blank(), 
        legend.position = "none")

fb_d <-
  LongData %>%
    filter(type == "Fire behavior", 
           value <= 600) %>%
    ggplot(aes(x = location, y= value,
               fill = location)) + theme_bw(14) +
      geom_violin() +
    labs(y = NULL, x = NULL, 
         title = "Fire behavior") + 
    facet_wrap( ~ var, scales = 'free') +
    scale_fill_manual(name = "Location", 
                      values = wes_palette("Cavalcanti1")[4:5]) + 
    theme(axis.text.x = element_blank(), 
          panel.grid.major.x = element_blank(), 
          legend.position = "bottom")

# Create list of panels
  gl <- lst(wx_d, fuel_d, fb_d)

  grid.arrange(
    grobs = gl,
    widths = c(1),
    heights = c(0.35, 0.25, 0.35)  )
```

## Variables associated with fire behavior

```{r glmer_mods, eval=TRUE}
# Mixed-effect regression models 
#
# Rate of spread
#   
  # Fit model 
    ros_imp <- with(imp_sc, suppressMessages(
                lme4::glmer(ros ~ RH + WindSpeed + 
                              FuelMoisture + tHa + 
                              (1|location/block/year/plot), 
                            family=Gamma(link = "log"), 
                            control=lme4::glmerControl(optimizer="bobyqa", 
                                                       optCtrl=list(maxfun=100000)) )) )
    ros_imp_pooled <- pool(ros_imp)$pooled  %>%
                          as_tibble()
    # Get terms
    ros_terms <- 
      full_join(
      summary(pool(ros_imp)) %>% 
        as_tibble() %>%
        rownames_to_column("row"),

      confint.mipo(pool(ros_imp)) %>%
        as_tibble() %>%
        rownames_to_column("row") )

#
# Maximum canopy temperature  
#
# Fit model 
canopy_imp <- with(imp_sc, suppressMessages(
              lme4::glmer(MaxC ~ RH + WindSpeed +  
                            FuelMoisture + tHa + 
                            (1|location/block/year/plot), 
                          family=Gamma(link = "log"), 
                          control=lme4::glmerControl(optimizer="bobyqa", 
                                                     optCtrl=list(maxfun=100000)) )) )

  canopy_imp_pooled <- pool(canopy_imp)$pooled  %>%
                      as_tibble()

  canopy_terms <- 
    full_join(
      summary(pool(canopy_imp)) %>% 
        as_tibble() %>%
        rownames_to_column("row"),

      confint.mipo(pool(canopy_imp)) %>%
        as_tibble() %>%
        rownames_to_column("row") )

#
# Maximum soil surface temperature  
#
# Fit model 
  soil_imp <- with(imp_sc, suppressMessages(
                  lme4::glmer(log(SoilMaxC+1) ~ RH + WindSpeed +  
                      FuelMoisture + tHa + 
                      (1|location/block/year/plot), 
                    family=Gamma(link = "log"), 
                    control=lme4::glmerControl(optimizer="bobyqa", 
                                               optCtrl=list(maxfun=100000)) )) )

soil_imp_pooled <- pool(soil_imp)$pooled  %>%
  as_tibble()

soil_terms <- 
  full_join(
    summary(pool(soil_imp)) %>% 
      as_tibble() %>%
      rownames_to_column("row"),

    confint.mipo(pool(soil_imp)) %>%
      as_tibble() %>%
      rownames_to_column("row") )

response_CIs <-   
  bind_rows(
    mutate(soil_terms, response = 'Temp. at surface'), 
    mutate(canopy_terms, response = 'Temp. above surface'),
    mutate(ros_terms, response = "Rate of spread") 
  ) %>%
  as_tibble() %>%
  rename(lwr = `2.5 %`, upr = `97.5 %`) %>%
  filter(term != '(Intercept)') %>%
  mutate(term = recode(term, tHa = 'Fuel load', 
                             WindSpeed = 'Wind speed', 
                             RH = 'Relative humidity', 
                             FuelMoisture = 'Fuel moisture')) 
```

### Rate of spread

```{r }
response_CIs %>%
  filter(response == "Rate of spread") %>%
  rename(t = statistic, 
         P = p.value) %>%
    select(term, t, P) %>%
  mutate(across(t:P, ~round(., 2))) %>%
  pander::pander("Summary statistics for Rate of spread.") 
```

### Aboveground temperatures

```{r }
response_CIs %>%
  filter(response == "Temp. above surface") %>%
  rename(t = statistic, 
         P = p.value) %>%
    select(term, t, P) %>%
  mutate(across(t:P, ~round(., 2))) %>%
  pander::pander("Summary statistics for flame temperature 15 cm above soil surface.") 
```

### Soil surface temperature

```{r }
response_CIs %>%
  filter(response == "Temp. at surface") %>%
  rename(t = statistic, 
         P = p.value) %>%
    select(term, t, P) %>%
  mutate(across(t:P, ~round(., 2))) %>%
  pander::pander("Summary statistics for soil surface temperature.") 
```

```{r CI_gg, fig.width=8, fig.height = 3.75}
# Plot confidence intervals
response_CIs %>%
  mutate(response =case_when(
                            response == "Temp. above surface" ~ "°C, canopy", 
                            response == "Temp. at surface" ~ "°C, soil", 
                            TRUE ~ response              )
                          ) %>%
  ggplot(aes(x = reorder(term, abs(estimate), max))) + theme_bw(16) + 
    geom_hline(yintercept = 0, color="black", linetype = 2) +
    geom_errorbar(aes(ymin = lwr, ymax = upr), 
                  size = 1.1, width = 0.2, 
                  color = "blue") +
    geom_point(aes(y = estimate), pch = 21, 
               stroke = 1.5, size = 4,
               color = "blue", 
               fill="lightblue") +
    labs(x = '', 
         y = 'Regression coefficient with 95% CI') + 
    coord_flip(ylim=c(-2,2.4)) + 
    facet_wrap(~ response) + 
    theme(axis.text.y = element_text(color="black", size = 14), 
          strip.text = element_text(size = 14), 
          panel.grid.major.y = element_blank(), 
          panel.grid.minor.x = element_blank())
```

## Multivariate analysis

```{r fit_pca}
# Fire behavior PCA 
  fb_d <- 
    imp_raw %>%
    select(MaxC, ros, SoilMaxC)

  fb_pca <- rda(fb_d ~ 1, 'euc', scale = T)
```

```{r pca_eig}
 summary(fb_pca)$cont$importance %>% pander::pander("Eigenvalues and proportion explained by composite axes (principal components).")
```

```{r}
# Testing for differences between locations 
pca_fc <-  envfit( fb_pca ~ location, imp_raw, 
                   choices = c(1:3), 
                   strata = imp_raw$year, 
                   199)$factors
pca_fc
```

No difference in fire behavior patterns between Hettinger and Central Grasslands; this is expected.

```{r envfit, eval=TRUE}
# Testing fire weather against PCA
pca_ev <- envfit(fb_pca ~ MaxWindSpeed+AirTemp+dpC+RH+VPD, 
                 data = imp_raw, 
                 choices = c(1:3), 
                 strata = imp_raw$location)
pca_ev
```

dpC, RH, and VPD had statistically-significant associations with variation in fire behavior, but even these relationships were weak.

```{r pca_dat}  
  # get ord data for ggplotting
  pca_gg <- gg_ordiplot(fb_pca, groups = imp_raw$location, 
                        plot=FALSE)

  pca_spp <- 
    scores(fb_pca, display = "species") %>%
    as.data.frame %>%
    as_tibble(rownames="response") %>%
    mutate(response = recode(response, 
                             "MaxC" = "canopy °C", 
                             "SoilMaxC" = 'soil °C', 
                             "ros" = "ROS"))

  # Vector 
  pca_vd <- scores(pca_ev,  "vectors") %>%
                as.data.frame %>% 
                rownames_to_column("gradient") %>%
              filter(gradient %in% c("dpC", "RH", "VPD")) %>%
              select(-PC3) %>%
            as_tibble( ) %>%
              mutate(gradient = ifelse(gradient == "dpC", 
                                       "DP", gradient))

  pca_scores <- lst(species=pca_spp, 
                    vectors=pca_vd)
  pca_scores$spiders <-
    pca_gg$df_spiders %>% 
    rename(PC1 = x, PC2 = y) %>%
    as_tibble()
```

```{r pca_gg, fig.height=4, fig.width=6, fig.cap='Principal Components Analysis of fire behavior data (response variables in red) for prescribed burns on rangeland at Hettinger (H) and Central Grasslands  (CG) Research Extension Centers. No difference between locations and only the vapor pressure deficit (VPD) and relative humidty (RH) gradients had significant associations with variation in fire behavior.' }
# ggplot PCA
  ggplot() + theme_bw(16) + 
    labs(x="PC 1", 
         y="PC 2") + 
    geom_vline(xintercept = 0, lty=3, color="darkgrey") +
    geom_hline(yintercept = 0, lty=3, color="darkgrey") +
    geom_point(data=pca_scores$spiders, 
                        aes(x=PC1, y=PC2, 
                            shape=Group, 
                            colour=Group), 
                        size=2)  +   
    geom_segment(data=pca_scores$spiders, 
                 aes(x=cntr.x, y=cntr.y,
                     xend=PC1, yend=PC2, 
                     color=Group), 
                 size=1.2, 
                 alpha = 0.75) +
    geom_point(data=pca_scores$spiders, 
               aes(x=PC1, y=PC2, fill=Group), 
               colour="grey40", 
               pch=21, 
               size=3, 
               stroke=2, alpha = 0.75) +
    geom_segment(data=pca_scores$vectors, 
                 aes(x=0, y=0, 
                     xend=PC1*1.6, 
                     yend=PC2*1.6),
                 arrow=arrow(length = unit(0.03, "npc")),
                 lwd=1.5) +
  geom_label(data=pca_scores$spiders,
               aes(x=cntr.x,
                   y=cntr.y,
                   label=Group,
                   color=Group),
               fontface="bold", size=4,
               label.size = 0,
               label.r = unit(0.5, "lines")) +
    geom_text(data=pca_scores$vectors,
              aes(x=PC1*1.75, 
                  y=PC2*1.75, 
                  label=gradient), 
              nudge_x = 0.06, 
              nudge_y =-0.05, 
              size=6, fontface="bold") + 
    geom_label(data=pca_scores$species, 
                aes(x=PC1/3, 
                    y=PC2/3, 
                    label=response), 
                label.padding=unit(0.25,"lines"),
                label.size = 0.5, 
                fontface="bold", 
               color="darkblue") +
    scale_color_manual(values = wes_palette("Cavalcanti1")[4:5]) + 
    scale_fill_manual(values = wes_palette("Cavalcanti1")[4:5])  +
    theme(panel.grid=element_blank(), 
          legend.position="none")

```

It is interesting that the gradient in temperature at the soil surface (soil °C) is orthogonal to both canopy temperature and rate of spread (canopy °C and ROS, respectively). 
This confirms trends observed in the regression analysis that the conditions most conducive to fast, high-intensity fires are not the same that result in surface heating.

Managers might be able to use high windspeeds to mitigate soil heating when fuels are dry, or focus on conditions with low RH to ensure litter burns down to the soil surface.

# Script

```{r script, ref.label=knitr::all_labels(!label %in% c('setup', 'dist_gg', 'data_summary_gg', 'CI_gg', 'pca_eig', 'pca_dat', 'pca_gg') ), echo=TRUE, eval=FALSE}
```