### 2022-01-25 paper/SupplementaryInfo.Rmd

---
title: Weather and fuel as modulators of grassland fire behavior in the northern Great Plains
subtitle: Supplementary information
date: '`r Sys.Date()`'
author: DA McGranahan, ME Zopfi, KA Yurkonis
header-includes:
  - \usepackage{amsmath}
  - \renewcommand{\familydefault}{\sfdefault}
  - \usepackage{hyperref}
output:
  pdf_document:
    highlight: tango
---

```{r setup, echo=FALSE, warning=FALSE, message = FALSE, results='hide'}
knitr::opts_chunk$set(message = FALSE, warning=FALSE, 
                      echo=FALSE, eval=FALSE)
```

# Study locations

```{r map, eval = TRUE, out.width='100%', fig.align='center', fig.cap='Main map: Study locations (triangles) within two EPA level 3 ecoregions in North Dakota. Inset: State of North Dakota (dark blue) within the Great Plains (light blue) with respect to the continental United States. \\label{fig:map}'}
knitr::include_graphics('figures/map.png')
```
\clearpage

# Data collection

For even more information about the FeatherFlame datalogger system, please see:

* McGranahan DA (2021) FeatherFlame: An Arduino-based thermocouple datalogging system to record wildland fire flame temperatures \emph{in agris}. Rangeland Ecology \& Management 76, 43–47 [DOI: 10.1016/j.rama.2021.01.008](https://doi.org/10.1016/j.rama.2021.01.008)

* [diyfirescience.info](https://diyfirescience.info)

```{r scheme, eval=TRUE, out.width='100%', fig.align='center', fig.cap='Schematic representation of the Sierpinski Triangle used to deploy 27 thermocouples across 9, 1 m equilateral triangles. Total plot area = 0.433 ha. \\label{fig:triangle}'}
knitr::include_graphics('figures/SierpinskiTriangle.png')
```

Simard et al. (1982) describe how rate of spread $r$ through an equilateral triangle with sides of length $D$ can be determined from the arrival times of the flame front at each point in the triangle sequentially\textemdash $t_1$, $t_2$, and $t_3$.

If $t_1 \neq t_2$,

\begin{equation} \label{eq:simard1}
\theta = tan\textsuperscript{-1} \left( \dfrac{2t_3 - t_2 - t_1}{\sqrt{3}\cdot (t_2 - t_1)}\right) 
\end{equation}

and rate of spread $r$ is

\begin{equation} \label{eq:simard2}
r = \dfrac{D\cdot cos\theta}{t_2 - t_1}
\end{equation}

otherwise, if $t_1 = t_2$, $\theta = 90$ and rate of spread $r$ is

\begin{equation} \label{eq:simard3}
r = D (\sqrt{3}/2)/({t_3 - t_1})
\end{equation}

\clearpage

```{r example_script}
### S E T U P 
##
# Additional packages required for analysis
  pacman::p_load(tidyverse, readr, mice, broom.mixed, vegan, lubridate)
# Additional script available via GitHub
  source('https://raw.githubusercontent.com/cran/mice/master/R/mipo.R')
#
##
### D A T A   P R E P A R A T I O N
##
# Load raw data directly from GitHub
  fp = 'https://raw.githubusercontent.com/devanmcg/SpatialFireBehavior/main'
#
# Data wrangling
#
 AllData <-  
    read_csv(paste0(fp, "/data/fromMZ/CompiledData2.csv")) %>%
    filter(location != "OAK") %>%
      mutate(date = as.Date(date, format = "%m/%d/%Y"),
               L = str_remove(location, "REC"), 
               B = str_sub(block, 1,3), 
               Ps = str_replace(pasture, "[.]", ""), 
               Ps = str_sub(Ps, 1,2), 
               patch = str_replace(patch, "[.]", ""),
               y = format(date, "%y")) %>%
        unite("FireCode", c(L,B,Ps,patch,y), sep=".") %>%
    mutate(time = str_remove(MaxTempTime, "[.]+[0-9]"))%>%
    unite(timestamp, c(date, time), sep = " ") %>%
    mutate(timestamp = as.POSIXct(timestamp, format = "%Y-%m-%d %H:%M:%S")) %>%
        select(FireCode, timestamp, plot, array, TC, MaxC, 
               AirTemp, RH, dpC, WindSpeed, 
               LAI, FMC, KgHa) 
# Isolate soil surface temperature (TC 4)
  SoilTemp <-
    filter(AllData, TC == 4) %>%
      select(FireCode, plot, array, MaxC) %>%
        rename(SoilC = MaxC)
# Summarize array-level data 
  DataMeans <- 
    AllData %>%
      filter(TC %in% c('1', '2', '3')) %>% 
      select(-timestamp) %>%
      pivot_longer(cols = c(MaxC:KgHa), 
                   names_to = "var",
                   values_to = "value") %>%
      group_by(FireCode, plot, array, var) %>%
        summarize(Mean = mean(value) ) %>%
      ungroup() %>%
      pivot_wider(names_from = var, 
                  values_from = Mean) 
# Calculate Vapor Pressure Deficit
  DataMeans <- 
    DataMeans %>%
      mutate(e  = 6.11*(10^((7.5*dpC)/(237.3+dpC))), 
             es = 6.11*(10^((7.5*AirTemp)/(237.3+AirTemp))), 
             VPD = es - e) %>%
      select(-e, -es)
# Calculate rate of spread by arrival time of flame front at sensors 
  D = 1   # Distance between thermocouples (m)
  ROS <- 
    AllData %>%
      filter(TC %in% c('1', '2', '3')) %>%
      mutate(timestamp = format(timestamp, "%H:%M:%S"), 
             ArrivalTime = seconds(hms(timestamp)) ) %>%
    select(FireCode, plot, array, ArrivalTime) %>%
    group_by(FireCode, plot, array) %>%
    arrange(ArrivalTime, .by_group = TRUE) %>% 
    mutate(position = order(order(ArrivalTime, decreasing=FALSE)), 
           position = recode(position, "1"="a", "2"="b", "3"="c"), 
           ArrivalTime = as.numeric(ArrivalTime) /60 ) %>%
    spread(position, ArrivalTime)  %>%
    ungroup %>% 
    # Apply equations from Simard et al. (1984)
    mutate( theta_rad = atan((2*c - b - a) / (sqrt(3)*(b - a))), 
            ros = case_when(
              a == b ~ (sqrt(3) / 2) / (c - a) , 
              a != b ~  (D*cos(theta_rad) / (b - a) ) 
            )) %>%
    select(-a, -b, -c, -theta_rad)
#
# Create final tibble for analysis 
#
  AnalysisData  <- 
    full_join(DataMeans, ROS) %>%
              left_join(SoilTemp) %>%
      filter( ros <= 40,       # remove outliers
              MaxC >= 40) %>%  # ditto
      rename(FuelMoisture = FMC, 
             SoilMaxC = SoilC) %>%
        mutate(FuelMoisture = ifelse(FuelMoisture >= 0, 
                                      FuelMoisture, NA), 
               FuelMoisture = FuelMoisture * 100) %>%
      separate(FireCode, into = c("location", "block", "pasture", 
                           "patch", "year"), 
               remove = F)
#
# Imputing missing values with mice package
#
  # Calculate imputed datasets on scaled data
    imp_sc <- AnalysisData %>% 
            select(-LAI, -JD) %>%
            mutate(ros = ifelse(ros >= 12, NA, ros),    # remove outliers
                   tHa = ifelse(tHa >= 4, NA, tHa)) %>% # ditto
            mutate_at(vars(AirTemp:tHa), ~as.numeric(scale(., center=F))) %>%
                      mutate(across(location:array, as.factor)) %>%
                       mice(m=50, seed = 23109, print=F)
#
##
###  S T A T I S T I C A L   M O D E L S
##
#
# Mixed-effect regression models on imputed datasets
#
# Rate of spread
#   
  # Fit model 
    ros_RH <- 
      with(imp_sc, suppressMessages(
            lme4::glmer(ros ~ RH + tHa +
                              FuelMoisture + WindSpeed +
                              (1|location/block/year/plot), 
                        family=Gamma(link = "log"), 
                        control=lme4::glmerControl(optimizer="bobyqa", 
                                      optCtrl=list(maxfun=100000)) )) )
  # Get terms
    ros_terms <- 
      full_join(
      summary(pool(ros_RH)) %>% 
        as_tibble() %>%
        rownames_to_column("row"), 
      confint.mipo(pool(ros_RH)) %>%
        as_tibble() %>%
        rownames_to_column("row") ) 
#    
# Maximum canopy temperature  
#
  # Fit model
    canopy_RH <- 
      with(imp_sc, suppressMessages(
            lme4::glmer(MaxC ~ RH + tHa +
                            FuelMoisture + WindSpeed +
                          (1|location/block/year/plot), 
                        family=Gamma(link = "log"), 
                        control=lme4::glmerControl(optimizer="bobyqa", 
                                      optCtrl=list(maxfun=100000)) )) )
  # Get terms
    canopy_terms <- 
      full_join(
        summary(pool(canopy_RH)) %>% 
          as_tibble() %>%
          rownames_to_column("row"), 
        confint.mipo(pool(canopy_RH)) %>%
          as_tibble() %>%
          rownames_to_column("row") ) 
#
# Maximum soil surface temperature  
#
  # Fit model 
    soil_RH <- 
      with(imp_sc, suppressMessages(
            lme4::glmer(log(SoilMaxC+1) ~ RH + tHa +
                        FuelMoisture + WindSpeed +
                        (1|location/block/year/plot), 
              family=Gamma(link = "log"), 
              control=lme4::glmerControl(optimizer="bobyqa", 
                            optCtrl=list(maxfun=100000)) )) )
  # Get terms
    soil_terms <- 
      full_join(
        summary(pool(soil_RH)) %>% 
          as_tibble() %>%
          rownames_to_column("row"), 
        confint.mipo(pool(soil_RH)) %>%
          as_tibble() %>%
          rownames_to_column("row") ) 
#
# Multivariate analysis
#
  # Reduce mids object to tibble   
    imp_raw <- 
      complete(imp_sc, 'long') %>%
      as_tibble() %>%
      unite("TreeID", c(location, block, pasture,
                        year, plot, array), sep = ".") %>%
      select(-patch, -.id, -.imp, -FireCode) %>%
      pivot_longer(names_to = "response", 
                   values_to = "values", 
                   -TreeID) %>%
      group_by(TreeID, response) %>%
      summarize(value = median(values)) %>%
      ungroup() %>%
      pivot_wider(names_from = response, 
                  values_from = value) %>%
      separate(TreeID, c("location", "block", "pasture",
                         "year", "plot", "array")) %>% 
      mutate(across(location:array, as.factor))
  # Fire behavior PCA 
    fb_d <- 
      imp_raw %>%
      select(MaxC, ros, SoilMaxC) 
    fb_pca <- rda(fb_d ~ 1, 'euc', scale = T)
  # Test differences between locations 
    envfit(fb_pca ~ location, imp_raw, 
           choices = c(1:2), 
           strata = imp_raw$year, 
           199)$factors
# Test fire weather against PCA
   envfit(fb_pca ~ MaxWindSpeed+AirTemp+dpC+RH+VPD, 
          data = imp_raw, 
          choices = c(1:3), 
          strata = imp_raw$location)
```

# Script

```{r script, ref.label=knitr::all_labels(label %in% c('example_script') ), echo=TRUE, eval=FALSE}
```