### 2023-10-12 paper/springer/CoverLetter/CoverLetter_R2.tex
surface update

\documentclass[parskip=half, 
			   fontsize=11pt,
			   paper=a4]				
{scrartcl}
\input{SetupOptions} % contains scrletter

\newcommand{\journal}{\emph{Environmental Management}}

%\vspace{-5em}

\begin{document}
	\begin{letter}{Editor\\
		\journal }
\setlength{\parindent}{10pt}

\opening{Dear Sir or Madam,}

We appreciate the opportunity to revise our submission for publication in \journal.
We've taken the editor's comments about further review seriously. 
We trust the editor will find that each of the second reviewer's comments have been fully replied to in this second round of review. 
Unfortunately the new major revisions period of 14 days is prohibitively short, and our attention in the previous review was focused on the detailed critique of the first reviewer. 
In this latest round, we have attended to all of the second reviewer's points and appreciate the additional detail the reviewer provided.

We look forward to working with the staff of \journal~to ensure our manuscript meets the standards of the journal.

\vspace{-3em} 
\closing{} %eg. Regards
%\cc{another dude}
\end{letter}

\end{document}